Design and Implementation of\\Anglican Probabilistic Programming Language

### 2016-09-02 slides/slides.tex
seems to becoming ready

\begin{frame}{More examples}
    \begin{itemize}
        \item \textit{Intruder detection} --- 
            given a log of \textbf{times} and \textbf{amounts} of
            payments in a bank account, how likely that the baccount was
            compromised?
        \pause
        \item \textit{Counterfactual reasoning} --- There are
            \textbf{two routes} from Jerusalem to Tel Aviv:
            1 and 443.  Based on traffic reports, I chose route
            1 and was late. Would I arrive on time If I chose 443
            instead?
        \pause 
        \item \textit{(Due to Stuart Russell)} If you observe that a
            student GPA is exactly 4.0 in a model of
            transcripts of students from the USA (GPA's from
            0.0 to 4.0) and India (GPA's from 0.0 to
            10.0) what is the probability that the student
            is from India? 
    \end{itemize}
\end{frame}

@@

\begin{frame}{Design challenges and choices}
\begin{itemize}
    \item Anglican syntax: Clojure + probabilistic constructs.
		\begin{itemize}
			\item Subset of Clojure.
			\item Special forms \texttt{sample} and \texttt{observe}.
			\item Macros to delimit Anglican code within a Clojure program.
		\end{itemize}
		\pause
		\bigskip
    \item Source-to-source compilation of Anglican into Clojure.
		\begin{itemize}
			\item CPS Transformation, with some tricks.
			\item Transparent use of Clojure functions from Anglican.
		\end{itemize}
		\pause
		\bigskip
    \item Inference algorithms.
		\begin{itemize}
            \item Required to run an Anglican program.
			\item Step in at checkpoints (\texttt{sample} and \texttt{observe}).
			\item Execute programs by calling continuations repeatedly.
		\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{Macro-based compilation}
    Anglican code macro-compiled into Clojure:
    \begin{minipage}{0.48\textwidth}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
(loop [data data]
  (if (seq data)
    (let [[x & data] data]
      (observe (gamma a b) x)
      (recur data))
    (predict :a a)))
\end{minted}
    \end{minipage}
    \begin{minipage}{0.44\textwidth}
    \vspace{2em}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
(fn loop [C23151 state data]
   (if (seq data)
     (let [[x & data] data]
       (->observe 'O23153 (gamma a b) x
         (fn do23152 [_ state]
           (loop C23151 state data))
         state))
     (fn []
       (C23151
         nil
         (add-predict state :a a)))))
\end{minted}
    \end{minipage}
\end{frame}

@@

@@

    \begin{minipage}{0.48\textwidth}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
    (fn [x y] (+ x y))
\end{minted}
    \end{minipage}
    \begin{minipage}{0.48\textwidth}
        \vspace{0.5em}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
    (fn [cont $state x y] 
      (fn []
        (cont (+ x y) $state)))
\end{minted}
      \vspace{0.5em}
    \end{minipage}

    ... and  called by \texttt{trampoline}:

>>>>

\begin{frame}{More examples}
    \begin{itemize}
        \item \textit{Intruder detection} --- 
            given a log of \textbf{times} and \textbf{amounts} of
            payments in a bank account, how likely that the baccount was
            compromised?
        \pause
        \item \textit{Counterfactual reasoning} --- There are
            \textbf{two routes} from Jerusalem to Tel Aviv:
            1 and 443.  Based on traffic reports, I chose route
            1 and was late. Would I arrive on time If I chose 443
            instead?
        \pause 
        \item \textit{(Due to Stuart Russell)} If you observe that a
            student GPA is exactly 4.0 in a model of
            transcripts of students from the USA (GPA's from
            0.0 to 4.0) and India (GPA's from 0.0 to
            10.0), what is the probability that the student
            is from India? 
    \end{itemize}
\end{frame}

@@

\begin{frame}{Design challenges and choices}
\begin{itemize}
    \item Anglican syntax: Clojure + probabilistic constructs.
		\begin{itemize}
			\item Subset of Clojure.
			\item Special forms \texttt{sample} and \texttt{observe}.
			\item Macros to delimit Anglican code within a Clojure program.
		\end{itemize}
		\pause
		\bigskip
    \item Source-to-source compilation of Anglican into Clojure.
		\begin{itemize}
			\item CPS transformation, with some tricks.
			\item Transparent use of Clojure functions from Anglican.
		\end{itemize}
		\pause
		\bigskip
    \item Inference algorithms.
		\begin{itemize}
            \item Required to run an Anglican program.
			\item Step in at checkpoints (\texttt{sample} and \texttt{observe}).
			\item Execute programs by calling continuations repeatedly.
		\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{Macro-based compilation}
    Anglican code macro-compiled into Clojure:
    \begin{minipage}{0.48\textwidth}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
(loop [data data]
  (if (seq data)
    (let [[x & data] data]
      (observe (gamma a b) x)
      (recur data))
    (predict :a a)))
\end{minted}
    \end{minipage}
    \begin{minipage}{0.44\textwidth}
    \vspace{2em}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
(fn loop [C23151 $state data]
   (if (seq data)
     (let [[x & data] data]
       (->observe 'O23153 (gamma a b) x
         (fn do23152 [_ $state]
           (loop C23151 $state data))
         $state))
     (fn []
       (C23151
         nil
         (add-predict $state :a a)))))
\end{minted}
    \end{minipage}
\end{frame}

@@

    \begin{minipage}{0.48\textwidth}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
    (fn [x y] (+ x y))
\end{minted}
    \end{minipage}
    \begin{minipage}{0.48\textwidth}
        \vspace{0.5em}
\begin{minted}[linenos, fontsize=\scriptsize]{clojure}
    (fn [cont $state x y] 
      (fn []
        (cont (+ x y) $state)))
\end{minted}
      \vspace{0.5em}
    \end{minipage}
    \pause
    ... and  called by \texttt{trampoline}:

@@
